\documentclass[a4paper,10pt,openany]{article}
\usepackage[utf8]{inputenc}
%\usepackage[czech]{babel}
\usepackage{geometry}
\usepackage{tikz,multirow,listingsutf8,multicol,amsmath,amsfonts}
\lstset{language=haskell,showstringspaces=false,stringstyle=\color{orange},basicstyle=\ttfamily\small,keywordstyle=\color{blue!60!black}, morekeywords={Aggregable,SegAVL}}
\geometry{left=20mm,right=20mm,top=20mm,bottom=20mm}
\parindent=0mm
\parskip=0mm

\begin{document}
\begin{center}
\pagenumbering{arabic}
{\huge \textsc{Segment Trees Haskell Implementation}}\\
\vspace{10mm} {\large Jiří Škrobánek\footnote[1]{Faculty of Mathematics and Physics, Charles University, {\ttfamily jiri@skrobanek.cz}}}\\
\vspace{10mm}\today, Prague

\end{center}

\section*{Introduction}
This document serves as a reference for \texttt{SegmentTree} Haskell module. The module provides an implementation of segment trees usable with any type of keys which implements \texttt{Ord} and can aggregate any associative function of values.

On a tree with \textit{n} elements all editing operations take $\mathcal{O}(\log n)$ time, as well as range queries. Insertion and deletion is implemented using AVL-tree balancing algorithms, yielding the same $\mathcal{O}(\log n)$ time complexity.

\section*{Example Usage}

In this example we build a tree on pairs of integers, to do this we must first instantiate the aggregation, in this case aggregation is simply addition. 

Firstly we build a tree on integers from 1 to 10. We update the value at 5 by adding 2 to it, lastly we ask for the sum of interval 2 to 8.

\lstinputlisting{../example.hs}

\section*{Complete List of Functionality}
\begin{lstlisting}
class Aggregable v where
    aggregate :: (Maybe v) -> (Maybe v) -> (Maybe v)
\end{lstlisting}
Values must be an instance of this class.
\begin{lstlisting}
data SegAVL k v
| Node {key::k, value::v, lsub::SegAVL k v, rsub::SegAVL k v, height::Int, agg::Maybe v}
| Empty 
\end{lstlisting}
\subsection*{Functions}
\begin{itemize}
\item \begin{lstlisting}
segAVLBuildFromList :: (Ord k, Aggregable v) => [(k,v)] -> SegAVL k v
\end{lstlisting}
First argument: Sorted list of key-value pairs.

This function creates the segment tree from a sorted list of key-value pairs.
\item \begin{lstlisting}
segAVLInsertAndBalance :: (Ord k, Aggregable v) => SegAVL k v->k->v-> SegAVL k v
\end{lstlisting}
First argument: Tree to insert into

Second argument: Key

Third argument: Value

This function inserts the key-value pait into the tree, overwriting old values.
\item \begin{lstlisting}
segAVLSetValue :: (Ord k, Aggregable v) => SegAVL k v->k->v->(v->v->v)->SegAVL k v
\end{lstlisting}

First argument: Tree where to set value

Second argument: Key

Third argument: Value

Fourth argument: Function taking as arguments old value and the third argument, producing new value if Key already present.

Similar to the previous function, but takes an extra argument which handles the replacing of old value.
\item \begin{lstlisting}
segAVLDelete :: (Ord k, Aggregable v) => SegAVL k v -> k -> SegAVL k v
\end{lstlisting}

First argument: Tree where to delete

Second argument: Key to delete

This function removes the given key from the tree.
\item \begin{lstlisting}
segAVLRange :: (Ord k, Aggregable v) => SegAVL k v -> k -> k -> Maybe v
\end{lstlisting}

First argument: Tree to query

Second argument: Minimum of range

Third argument: Maximum of range

This function return the aggregated value from interval between the second and third parameter.
\item \begin{lstlisting}
segAVLFind :: (Ord k, Aggregable v) => SegAVL k v -> k -> Maybe v
\end{lstlisting}

First argument: Tree search

Second argument: Key to find

Returns the value associated with given key or Nothing.
\item \begin{lstlisting}
segAVLMember :: (Ord k, Aggregable v) => SegAVL k v -> k -> Bool
\end{lstlisting}

First argument: Tree to search

Second argument: Key to find

Return a Boolean determining whether the key is in the tree.
\item \begin{lstlisting}
segAVLToList :: (Ord k, Aggregable v) => SegAVL k v -> [(k,v)]
\end{lstlisting}

First argument: Tree to convert

Produces an ordered set of key-value pairs from the tree.
\item \begin{lstlisting}
segAVLGetMin :: SegAVL k v -> Maybe (k,v)
\end{lstlisting}

First argument: Tree to search

Returns the minimum key-value pair present in the tree.
\item \begin{lstlisting}
segAVLGetMax :: SegAVL k v -> Maybe (k,v)
\end{lstlisting}

First argument: Tree to search

Returns the maximum key-value pair present in the tree.
\item \begin{lstlisting}
segAVLLowerBound :: (Ord k) => SegAVL k v -> k -> Maybe (k,v)
\end{lstlisting}

First argument: Tree to search

Second argument: Inclusive lower bound

Return the key-value pair where the key is the smallest not lesser then the key provided.
\item \begin{lstlisting}
segAVLUpperBound :: (Ord k) => SegAVL k v -> k -> Maybe (k,v)
\end{lstlisting}

First argument: Tree to search

Second argument: Exclusive upper bound

Return the key-value pair where the key is the greatest lesser then the key provided.
\end{itemize}
\end{document}